\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-26 paper.qmd
Updated Figure 12 caption

To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the data structure present in $9\text{-}D$ (@fig-model-pbmc-author-proj (b)), we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data. The @fig-NLDR-variety (a) shows three well-separated clusters with big separations. However, as shown in @fig-model-pbmc-author-proj, there is no big separation between three clusters in $9\text{-}D$. Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset in $9\text{-}D$. But, when visualizing the *model-in-th-data-space* some unobserved structures can be seen. Some clusters have non-linear continuity patterns and high-density patches (@fig-model-pbmc-author-proj).

@@

```{r}
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "Model in \\gD{}, on the UMAP layout, and three views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) The full video is available at <https://youtu.be/VqqWuE0Jj6A>."

>>>>

To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the data structure present in $9\text{-}D$ (@fig-model-pbmc-author-proj (b)), we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data. The @fig-NLDR-variety (a) shows three well-separated clusters with big separations. However, as shown in @fig-model-pbmc-author-proj, there is no big separation between three clusters in $9\text{-}D$. Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset in $9\text{-}D$. But, when visualizing the *model-in-the-data-space* some unobserved structures can be seen. Some clusters have non-linear continuity patterns and high-density patches (@fig-model-pbmc-author-proj).

@@

```{r}
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "Model in \\gD{}, on the UMAP layout (a), and three views of the fit in projections from $9\\text{-}D$ (b, c, d), for the PBMC3k data ($a_1 = 0.046$, $b = 676 \\  (26, \ 26)$, $m = 92$, benchmark value to remove large edges is $0.08$). The three well-separated clusters observed in the \\gD{} layout of UMAP (@fig-NLDR-variety (a)) is also visible when visualizing the model overlaid on the data space. But the separation between the clusters are small in $9\\text{-}D$. The model shows non-linear clusters and densed points within clusters in $9\\text{-}D$ space, which are some additional patterns not visible in the \\gD{} representation (@fig-NLDR-variety (a)). The full video is available at <https://youtu.be/VqqWuE0Jj6A>."

### 2024-08-26 paper.tex
Updated Figure 12 caption

To determine whether the UMAP representation with the specific
(hyper-)parameter choice suggested by \citet{chen2023} preserves the
data structure present in \(9\text{-}D\)
(Figure~\ref{fig-model-pbmc-author-proj} (b)), we visualize the model
constructed with UMAP overlaid on the \(9\text{-}D\) data. The
Figure~\ref{fig-NLDR-variety} (a) shows three well-separated clusters
with big separations. However, as shown in
Figure~\ref{fig-model-pbmc-author-proj}, there is no big separation
between three clusters in \(9\text{-}D\). Therefore, the suggested UMAP
representation (Figure~\ref{fig-NLDR-variety} (a)) does not accurately
represent the structure of PBMC3k dataset in \(9\text{-}D\). But, when
visualizing the \emph{model-in-th-data-space} some unobserved structures
can be seen. Some clusters have non-linear continuity patterns and
high-density patches (Figure~\ref{fig-model-pbmc-author-proj}).

@@

\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the UMAP
layout, and three views of the fit in projections from \(9\text{-}D\),
for the PBMC3k data (\(b = 870   (30,  29)\), \(m = 135\), benchmark
value to remove large edges is \(0.099\)) The full video is available at
\url{https://youtu.be/VqqWuE0Jj6A}.}

>>>>

To determine whether the UMAP representation with the specific
(hyper-)parameter choice suggested by \citet{chen2023} preserves the
data structure present in \(9\text{-}D\)
(Figure~\ref{fig-model-pbmc-author-proj} (b)), we visualize the model
constructed with UMAP overlaid on the \(9\text{-}D\) data. The
Figure~\ref{fig-NLDR-variety} (a) shows three well-separated clusters
with big separations. However, as shown in
Figure~\ref{fig-model-pbmc-author-proj}, there is no big separation
between three clusters in \(9\text{-}D\). Therefore, the suggested UMAP
representation (Figure~\ref{fig-NLDR-variety} (a)) does not accurately
represent the structure of PBMC3k dataset in \(9\text{-}D\). But, when
visualizing the \emph{model-in-the-data-space} some unobserved
structures can be seen. Some clusters have non-linear continuity
patterns and high-density patches
(Figure~\ref{fig-model-pbmc-author-proj}).

@@

\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the UMAP
layout (a), and three views of the fit in projections from
\(9\text{-}D\) (b, c, d), for the PBMC3k data (\(a_1 = 0.046\),
\(b = 676 \  (26,  26)\), \(m = 92\), benchmark value to remove large
edges is \(0.08\)). The three well-separated clusters observed in the
\gD{} layout of UMAP (Figure~\ref{fig-NLDR-variety} (a)) is also visible
when visualizing the model overlaid on the data space. But the
separation between the clusters are small in \(9\text{-}D\). The model
shows non-linear clusters and densed points within clusters in
\(9\text{-}D\) space, which are some additional patterns not visible in
the \gD{} representation (Figure~\ref{fig-NLDR-variety} (a)). The full
video is available at \url{https://youtu.be/VqqWuE0Jj6A}.}
